---
title: "Title"
author: "Christian Oppegård Moen"
date: "DD MM YYYY"
output: 
  bookdown::pdf_document2:
    toc_depth: '3' # Depth of contents
    number_sections: false
  # pdf_document:
  # #   toc: no
  #   toc_depth: '3'
subtitle: Course
urlcolor: blue
editor_options: 
  chunk_output_type: console
header-includes:
- \usepackage[width=0.8\textwidth]{caption} # add caption to figures.
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  echo = T, tidy=T, message=F, warning=F,
  strip.white=F, # Strips whitespace
  prompt=F, #?
  cache=T, # Cache for quick knit.
  # root.dir = "./subfolder",
  size="scriptsize",
  fig.width=7, 
  fig.height=5, 
  fig.align = "center"
)
```

```{r config, include=F}
if (F){
  setwd("C:/Users/kikka/OneDrive - NTNU/Fysmat/8 Semester V2022/RomStat/romstat/project3")
  options(error=recover)
  options(error=NULL)
}
```

# Libraries & misc
```{r libraries, options}
library(Rmisc)
library(glmm)
library(ggplot2)
library(tidyr)
library(dplyr)
library(nlme)
library(lme4)
```

```{r defaultSettings, options}
defaultMar = c(5.1, 4.1, 4.1, 2.1)

```

# Data
Data format is of tab separated values regarding noise on campus. Some variables are shown in the printout below. The response is `spmTriv` and `spmEff` which are "trivsel" and "effektivitet", respectively. The response is evaluated for each building block.

## Load data and reformat
The time is on a unfeasible format, so we format it to total seconds used.
```{r reformatTime, options}
pathData = "./data"
d = read.delim("./data/data-315297-2023-03-08-1434-utf.txt", header=T)

# Reformat the time to be total time in seconds
formatTime <- function(t) {
  tSplit = strsplit(t, ' ')[[1]]
  s=0
  for (i in seq(1, length(tSplit), 2)){
    s = s + switch(
      tSplit[i+1],
      "dag" = strtoi(tSplit[i])*24*3600,
      "dager" = strtoi(tSplit[i])*24*3600,
      "time" = strtoi(tSplit[i])*3600,
      "timer" = strtoi(tSplit[i])*3600,
      "minutt" = strtoi(tSplit[i])*60,
      "minutter" = strtoi(tSplit[i])*60,
      "sekund" = strtoi(tSplit[i]),
      "sekunder" = strtoi(tSplit[i]),
      0
    )
  }
  return(s)
}

ftimes = unlist(lapply(d$Svartid, formatTime))
head(cbind(old=d$Svartid, new=ftimes))
d$Svartid=ftimes
```

```{r freeTextToFile, eval=F, include=F}
write.csv(d[d[,'spmTekst']!="",c('NR','spmTekst')],"./data/freeTxt.csv", row.names = FALSE)
```



# Initial values
In this section we take the initial look at the data by considering sample means, SDs, conf.ints and some graphs being plots of said values, histograms and more.

## "How many" numbers
How many did something about the noise.
```{r howManyTiltak, options}
# head(d)
# head(d[,18:(18+6)])
dTiltak = d[,18:(18+6)]
hasAns = function(x){return (sum(x!=""))}
dTiltakSum = apply(dTiltak, FUN = hasAns, MARGIN = 2)
dTiltakSum
```

## Total mean, SD, CI
We compute the sample mean, SD and confidence intervals of `effektivitet` and `trivsel` of the entire data set. We also take a look at their distribution by looking at the histograms.
```{r totMeanSDCI, options}
effTrivInit = cbind(t(apply(d[, c('spmEff', 'spmTriv')], 2, FUN = CI)), c(sd(d$spmEff), sd(d$spmTriv)))
colnames(effTrivInit) = c(colnames(effTrivInit)[1:3], 'SD')
effTrivInit

# Histograms
histCounts = as.data.frame(rbind(hist(d$spmTriv, breaks = seq(0.5,5.5,1), plot = F)$counts,
               hist(d$spmEff, breaks = seq(0.5,5.5,1), plot = F)$counts))
rownames(histCounts) = c('trivsel', 'effektivitet')
colnames(histCounts) = 1:5

histCounts

barplot(
  ((as.matrix(histCounts))), 
  col = c("lightblue", 'darkblue'),
  border = "white",
  # main="Trivsel",
  ylab="Antall svar",
  xlab = "P\u00E5virkning",
  beside = T,
  las=1
  # ylim = c(0,5)
  # space=0.1
)
legend(
  "topright",
  legend = c("Trivsel", "Effektivitet"),
  fill = c("lightblue", 'darkblue'), bty = 'n')
```


```{r save_totMeanSDCI_Hist, include=FALSE, eval=FALSE}
pdf(file = './figures/histTrivsel&EffektivitetTotal.pdf', width = 10,height = 6)

barplot(
  ((as.matrix(histCounts))), 
  col = c("lightblue", 'darkblue'),
  border = "white",
  # main="Trivsel",
  ylab="Antall svar",
  xlab = "P\u00E5virkning",
  beside = T,
  las=1
  # ylim = c(0,5)
  # space=0.1
)
legend(
  "topright",
  legend = c("Trivsel", "Effektivitet"),
  fill = c("lightblue", 'darkblue'), bty = 'n')
dev.off()
```


## Location Analysis
Let's look at some initial location based data. First, how many answers per building.
```{r perBuilInit, options}
# All locations/buildings
locs = unique(d$spmHvor) # all locations in this dataset
buildInit = data.frame(
  matrix(ncol=length(locs),
         nrow=5, 
         dimnames=list(c('count', 'meanEff', 'SDEff', 'meanTriv', 'SDTriv'), locs))
)

for (loc in locs){
  buildInit[loc] = c(sum(d$spmHvor==loc),
             mean(d$spmEff[d$spmHvor==loc]),
              sd(d$spmEff[d$spmHvor==loc]),
             mean(d$spmTriv[d$spmHvor==loc]),
              sd(d$spmTriv[d$spmHvor==loc]))
}

# Count, mean and standard deviation per building
round(buildInit,1)

# Sorted values for readability
sort(round(buildInit['meanTriv',],1))
sort(round(buildInit['meanEff',],1))
sort(round(buildInit['SDTriv',],1))
sort(round(buildInit['SDEff',],1))

# Barplot of counts for readability
buildNames = c(
  'Elektro D',
  "Elektro B",
  "Kjelhuset",
  "Elektro E",
  "Realfag",
  "Sentralbygg 2",
  "Verkstedteknisk",
  "Byggteknsik",
  "Sentralbygg 1",
  "Handelsh\u00F8gskolen",
  "Tapirbygget",
  "Berg",
  "Gamle elektro",
  "Hovedbygget",
  "Varmeteknisk",
  "Gruveteknisk",
  "Gamle fysikk",
  "IT-bygget syd",
  "Kjemiblokk 1",
  "Kjemiblokk 2",
  "Elektro A",
  "Kjemiblokk 5",
  "IPD"
)
par(mar=defaultMar+c(3,0,0,0))
barplot(
  ((as.matrix(buildInit['count',]))), 
  col = c("lightblue"),
  border = "white",
  # main="Trivsel",
  ylab="Antall svar",
  names.arg = buildNames,
  # xlab = "Lokasjon",
  beside = T,
  las=2
  # ylim = c(0,5)
  # space=0.1
)
par(mar=defaultMar)
```

```{r saveBuildNumber, include=F, eval=F}
pdf(file = './figures/histAnsPerBuild.pdf', width = 10,height = 6)
par(mar=defaultMar+c(3,0,0,0))
barplot(
  ((as.matrix(buildInit['count',]))), 
  col = c("lightblue"),
  border = "white",
  # main="Trivsel",
  ylab="Antall svar",
  # xlab = "Lokasjon",
  beside = T,
  las=2,
  names.arg = buildNames
  # ylim = c(0,5)
  # space=0.1
)
dev.off()
```


### Barplot of responses
Compute responses
```{r response, options}
# initiate per building empirical mean data frame
buildMeans = data.frame(
  matrix(ncol=length(locs),
         nrow=2, 
         dimnames=list(c('trivsel', 'effektivitet'), locs))
  )
# initiate per building standard deviation data frame
buildSd = data.frame(
  matrix(ncol=length(locs),
         nrow=2, 
         dimnames=list(c('trivsel', 'effektivitet'), locs))
  )
# Compute means and SDs
for (loc in locs){
  buildMeans[loc] = c(mean(d$spmTriv[d$spmHvor==loc]),
             mean(d$spmEff[d$spmHvor==loc]))
  buildSd[loc] = c(sd(d$spmTriv[d$spmHvor==loc]),
             sd(d$spmEff[d$spmHvor==loc]))
} 

# Mean values per location
round(sort(buildMeans['trivsel',]),2)
round(sort(buildMeans['effektivitet',]),2)
```

```{r writeTo.dat, eval=F, include=F}
# means
write.table(
  t(buildMeans['trivsel',]), 
  file= paste(pathData,'/trivselPerBygning.dat',sep = ""), 
  row.names = T, sep = ',', quote = F
)
write.table(
  t(buildMeans['effektivitet',]), 
  file= paste(pathData,'/effektivitetPerBygning.dat',sep = ""), 
  row.names = T, sep = ',', quote = F
)
# Standard deviations (sd)
write.table(
  t(buildSd['trivsel',]), 
  file= paste(pathData,'/trivselPerBygningSD.dat',sep = ""), 
  row.names = T, sep = ',', quote = F
)
write.table(
  t(buildSd['effektivitet',]), 
  file= paste(pathData,'/effektivitetPerBygningSD.dat',sep = ""), 
  row.names = T, sep = ',', quote = F
)
```

Plot responses
```{r barplotResponses, options}
# æ is \u00E6
# ø is \u00F8
# å is \u00E5
par(mar=defaultMar+c(3,0,0,0))
barplot(
  ((as.matrix(buildMeans))), 
  col = c("lightblue", 'darkblue'),
  border = "white",
  # main="Trivsel",
  ylab="P\u00E5virkning",
  beside = T,
  las=2,
  names.arg = buildNames,
  ylim = c(0,5)
  # space=0.1
)

legend(
  "top",
  legend = c("Trivsel", "Effektivitet"),
  fill = c("lightblue", 'darkblue'), bty = 'n')
d['Elektro_D.B2',]
head(d)
mTriv= max(buildMeans['trivsel',])
mEff = max(buildMeans['effektivitet',])
buildMeans['effektivitet',buildMeans['effektivitet',]==mEff]
buildMeans['trivsel',buildMeans['trivsel',]==mTriv]
```

```{r barplotResponseSD, options}
barplot(
  ((as.matrix(buildSd))), 
  col = c("lightblue", 'darkblue'),
  border = "white",
  # main="Trivsel",
  ylab="Standardavvik",
  beside = T,
  las=2,
  names.arg = buildNames
  # ylim = c(0,5)
  # space=0.1
)
legend(
  "top",
  legend = c("Trivsel", "Effektivitet"),
  fill = c("lightblue", 'darkblue'), bty = 'n')
```

```{r saveResponseBarplot, include=FALSE, eval=FALSE}
# æ is \u00E6
# ø is \u00F8
# å is \u00E5
pdf(file = './figures/histTrivsel&Effektivitet.pdf', width = 10,height = 6)
par(mar=defaultMar+c(3,0,0,0))
barplot(
  ((as.matrix(buildMeans))), 
  col = c("lightblue", 'darkblue'),
  border = "white",
  # main="Trivsel",
  ylab="P\u00E5virkning",
  beside = T,
  las=2,
  ylim = c(0,5),
  names.arg = buildNames,
  # space=0.1
)
legend(
  "top",
  legend = c("Trivsel", "Effektivitet"),
  fill = c("lightblue", 'darkblue'), bty = 'n')
dev.off()
```


```{r saveBarplotResponsesSD, eval=FALSE, include=FALSE}
pdf(file = './figures/histTrivsel&EffektivitetSD.pdf', width = 10,height = 6)
par(mar=defaultMar+c(3,0,0,0))
barplot(
  ((as.matrix(buildSd))), 
  col = c("lightblue", 'darkblue'),
  border = "white",
  # main="Trivsel",
  ylab="Standardavvik",
  beside = T,
  las=2,
  names.arg = buildNames
  # ylim = c(0,5)
  # space=0.1
)
legend(
  "top",
  legend = c("Trivsel", "Effektivitet"),
  fill = c("lightblue", 'darkblue'), bty = 'n')
dev.off()
par(mar=defaultMar)
```


<!-- # Make building data -->
<!-- ```{r buildingSpecificVariables, options} -->
<!-- # buildings = split(d, f=d$spmHvor) -->
<!-- # for (build in buildings){ -->
<!-- #   print(mean(build$spmTriv)) -->
<!-- # } -->
<!-- ``` -->

<!-- # GLMM fit on errytin' -->
<!-- ```{r GLMMall, eval=F} -->
<!-- # set.seed(420) -->
<!-- # fitAll = glmm(spmTriv~0+Svartid, varcomps.names = c(""), data = d, family.glmm = Gaussian) -->
<!-- # dSummary = summary(d) -->
<!-- # head(d) -->
<!-- ``` -->

<!-- # GLM fit on Svartid -->
<!-- ```{r GLMtid, options} -->
<!-- # # Fit -->
<!-- # fitTidTriv = glm(factor(spmTriv, seq(1,5,1))~Svartid , data = d, family = "binomial") -->
<!-- # fitTidEff = glm(factor(spmEff, seq(1,5,1))~Svartid , data = d, family = "binomial") -->
<!-- #  -->
<!-- # summary(fitTidTriv)$coefficients -->
<!-- # summary(fitTidEff)$coefficients -->
<!-- # summary(fitTidEff) -->
<!-- ``` -->

<!-- # GLM fit on noise types -->
<!-- ```{r glmNoiseTypes, options} -->
<!-- # factors = c('spmTriv', 'spmEff', -->
<!-- #             'byggestoy1', 'personstoy1','trafikk1','vifte1', 'annet1', -->
<!-- #             'byggestoy2', 'personstoy2','trafikk2','vifte2', 'annet2' -->
<!-- #             ) -->
<!-- # dFactored = d -->
<!-- # # factorize = function() -->
<!-- # for (f in factors){ -->
<!-- #   dFactored[f,] = factor(d[f,], seq(1,5,1), ordered = T) -->
<!-- # } -->
<!-- #  -->
<!-- # fitTrivSource = glm( -->
<!-- #   factor(spmTriv, seq(1,5,1), ordered = T)~byggestoy1 + personstoy1 + trafikk1 + vifte1 + annet1, -->
<!-- #   data = dFactored, family = "binomial" -->
<!-- # ) -->
<!-- # summary(fitTrivSource) -->
<!-- # cov(d$spmTriv, d$byggestoy1) -->
<!-- # cov(d$spmEff, d$byggestoy2) -->
<!-- # par(d) -->
<!-- ``` -->






<!-- # Linear models -->
<!-- Fitting linear model with covariates being noise types. -->
<!-- ```{r linMods, options} -->
<!-- lmTrivSource = lm( -->
<!--   spmTriv ~ -1 + byggestoy1 + personstoy1 + trafikk1 + vifte1 + annet1, -->
<!--   data=d -->
<!-- ) -->
<!-- lmNTcoefs = summary(lmTrivSource)$coefficients -->
<!-- signCoefs = lmNTcoefs[lmNTcoefs[,4]<1e-2, ] -->
<!-- lmNTcoefs # all coefficients -->
<!-- signCoefs # significant coefficients -->
<!-- ``` -->


<!-- ```{r numericVals, options} -->
<!-- dNumeric = select_if(d,is.numeric) -->
<!-- # head(dNumeric[,2:7]) -->
<!-- lmTrivSource = lm(spmTriv~., data = dNumeric[2:7]) -->
<!-- lmTrivCoefs = summary(lmTrivSource)$coefficients -->
<!-- lmTIntercept = lmTrivCoefs['(Intercept)', 'Estimate'] -->
<!-- lmTBygg = lmTrivCoefs['byggestoy1', 'Estimate'] -->
<!-- lmTPers = lmTrivCoefs['personstoy1', 'Estimate'] -->
<!-- lmTVift = lmTrivCoefs['vifte1', 'Estimate'] -->
<!-- yEst = lmTBygg*(1:5) + lmTPers*(1:5) + lmTVift*(1:5) -->
<!-- plot(dNumeric[,'byggestoy1'], dNumeric[,'spmTriv'], ) -->
<!-- plot(yEst) -->
<!-- ``` -->


# Pairs plots
Visualization of each data value compared to each other. Neat for initial data observations.
```{r pairs, options}
# pairs(dNumeric[,2:7])
library(GGally)
dNumeric = select_if(d,is.numeric)
ggpairs(dNumeric[,2:7]) # Trivsel pairs
ggpairs(dNumeric[,8:13]) # Effektivitet pairs
```

# Noise type analysis
In this section we will take a look at each noise type and how much people feel like it influences their efficiency and well-being. 
<!-- First we consider what each individual has set the noise type influence e.g., `byggestoy1`$=3$, and then we consider a weighted result where the wights are the overall influence divided by $5$, e.g., `effektivitet/5*byggestoy1`. -->

## Noise type mean and SD barplot. (Discarded)
```{r tempNecessities, options}
# setup
ntColnames = c('Konstruksjon', 'Person', 'Trafikk', 'Vifte', 'Annet')
nts = c('Konstruksjon', 'Person', 'Trafikk', 'Vifte', 'Annet')
ntTriv = colnames(d)[7:11]
ntEff = colnames(d)[13:17]
```

```{r noiseTypes, eval=T}


# compute mean and sd
ntDf = data.frame(cbind(
  nts=1:length(nts),
  meanEff=apply(d[, ntEff], FUN = mean, MARGIN = 2),
  sdEff=apply(d[, ntEff], FUN = sd, MARGIN = 2),
  meanTriv=apply(d[, ntTriv], FUN = mean, MARGIN = 2),
  sdTriv=apply(d[, ntTriv], FUN = sd, MARGIN = 2)
))

# compute mean and sd weighted
ntDfW = data.frame(cbind(
  nts=1:length(nts),
  meanEff=apply(d[, ntEff]*(d$spmEff/5), FUN = mean, MARGIN = 2),
  sdEff=apply(d[, ntEff]*(d$spmEff/5), FUN = sd, MARGIN = 2),
  meanTriv=apply(d[, ntTriv]*(d$spmEff/5), FUN = mean, MARGIN = 2),
  sdTriv=apply(d[, ntTriv]*(d$spmEff/5), FUN = sd, MARGIN = 2)
))

# Plot not weighted
ntDf_maxMean = max(ntDf[,c('meanEff','meanTriv')])
ntDf_maxSD = max(ntDf[,c('sdEff','sdTriv')])

ntDfLong <- ntDf |> 
  pivot_longer(cols = -nts,names_to = "Type") |>
  mutate(scaled_value=ifelse(
    Type %in% c('meanEff','meanTriv'),
    value,
    value/ntDf_maxSD*ntDf_maxMean))


# Plot Weighted
ntDfW_maxMean = max(ntDfW[,c('meanEff','meanTriv')])
ntDfW_maxSD = max(ntDfW[,c('sdEff','sdTriv')])

ntDfWLong <- ntDfW |> 
  pivot_longer(cols = -nts,names_to = "Type") |>
  mutate(scaled_value=ifelse(
    Type %in% c('meanEff','meanTriv'),
    value,
    value/ntDfW_maxSD*ntDfW_maxMean))

ggplot(ntDfWLong, aes(x=nts, y = scaled_value,fill= Type)) +
  geom_col(position="dodge") +
  scale_y_continuous(sec.axis = sec_axis(~ . /ntDf_maxMean*ntDf_maxSD , name = "Standardavvik"))+
  scale_fill_manual(
    values = c('#59f7ea', '#be77f9', "#ccfcf9", '#dbbcf5'),
    labels = c('Effektivitet', 'Trivsel', 'SD Effektivitet', 'SD Trivsel')
  ) +
  labs(y="P\u00E5virkning", x="St\u00F8ytype") +
  scale_x_discrete(limit = nts)+
  theme_minimal() +
  theme(legend.title = element_blank())

```

```{r saveNoiseTypes, eval=F, include=F}
pdf(file = './figures/noiseTypeBar.pdf', width = 6,height = 6)
ggplot(ntDfLong,aes(x=nts, y = scaled_value,fill= Type)) + 
  geom_col(position="dodge") + 
  scale_y_continuous(sec.axis = sec_axis(~ . /ntDf_maxMean*ntDf_maxSD , name = "Standardavvik"))+
  scale_fill_manual(
    values = c('#59f7ea', '#be77f9', "#ccfcf9", '#dbbcf5'),
    labels = c('Effektivitet', 'Trivsel', 'SD Effektivitet', 'SD Trivsel')
  )+
  labs(y="P\u00E5virkning", x="St\u00F8ytype") + 
  scale_x_discrete(limit = nts)+
  theme_minimal() + 
  theme(legend.title = element_blank())
dev.off()
```

```{r removed, include=FALSE, eval=FALSE}

# æ is \u00E6
# ø is \u00F8
# å is \u00E5
# ggplot(ntDfLong,aes(x=nts, y = scaled_value,fill= Type)) + 
#   geom_col(position="dodge") + 
#   scale_y_continuous(sec.axis = sec_axis(~ . /ntDf_maxMean*ntDf_maxSD , name = "Standardavvik"))+
#   scale_fill_manual(
#     values = c('#59f7ea', '#be77f9', "#ccfcf9", '#dbbcf5'),
#     labels = c('Effektivitet', 'Trivsel', 'SD Effektivitet', 'SD Trivsel')
#   ) +
#   labs(y="P\u00E5virkning", x="St\u00F8ytype") + 
#   scale_x_discrete(limit = nts)+
#   theme_minimal() + 
#   theme(legend.title = element_blank())

# With conf int bars
# ntDfMean = ntDf[, c('nts','meanEff', 'meanTriv')]
# ntDfSd = ntDf[, c('nts','sdEff', 'sdTriv')]
# ntDfMeanLong <- ntDfMean |> 
#   pivot_longer(cols = -nts,names_to = "Type")
# ntDfSdLong <- ntDfSd |> 
#   pivot_longer(cols = -nts,names_to = "Type")
# ntDfMeanLong$sd = ntDfSdLong$value

# ggplot(ntDfMeanLong, aes(x=nts, y = value,fill= Type)) + 
#   geom_col(position="dodge") + 
#   # scale_y_continuous(sec.axis = sec_axis(~ . /ntDf_maxMean*ntDf_maxSD , name = "Standardavvik"))+
#   scale_fill_manual(
#     values = c('#59f7ea', '#be77f9'),
#     labels = c('Effektivitet', 'Trivsel')
#   )+
#   geom_errorbar(
#     aes(
#       # x=nts,
#       ymin = value - 1.96*sd,
#       ymax = value + 1.96*sd),
#       color = "black",
#     position=position_dodge(.9),
#     width=.2
#   ) +
#   # facet_grid(cols = vars(Type))+
#   labs(y="P\u00E5virkning", x="St\u00F8ytype") + 
#   scale_x_discrete(limit = nts)+
#   theme_minimal() + 
#   theme(legend.title = element_blank())
```

# Noise types with confidence intervals
```{r ntConfInt, options}
# Compute mean and confidence intervals
ntDfCi = data.frame(
  cbind(
    nts=1:length(nts),
    t(apply(d[,ntEff], FUN = CI, MARGIN=2)),
    t(apply(d[,ntTriv], FUN = CI, MARGIN=2))
    )
)

# colnames(ntDfCi)
meansTemp = c('nts','mean','mean.1')
upperTemp = c('nts','upper','upper.1')
lowerTemp = c('nts','lower','lower.1')
dfMTemp = ntDfCi[,meansTemp]
dfUTemp = ntDfCi[,upperTemp]
dfLTemp = ntDfCi[,lowerTemp]

ntDfCiLong <- dfMTemp|>
  pivot_longer(cols=-nts, names_to='Type')
ntDfULong = dfUTemp |>
  pivot_longer(cols=-nts, names_to='Type')
ntDfLLong = dfLTemp |>
  pivot_longer(cols=-nts, names_to='Type')

ntDfCiLong$upper=ntDfULong$value
ntDfCiLong$lower=ntDfLLong$value

ntDfCiLong$nts=nts[ntDfCiLong$nts]
# rep(c('eff','triv'), as.integer(length(nts)))
ntDfCiLong$Type=rep(c('eff','triv'), as.integer(length(nts)))
colnames(ntDfCiLong) = c('nts', 'Type', 'mean', 'upper', 'lower')

ggplot(ntDfCiLong, aes(x=nts, y = mean ,fill= Type)) + 
  geom_col(position="dodge") + 
  # scale_y_continuous(sec.axis = sec_axis(~ . /ntDf_maxMean*ntDf_maxSD , name = "Standardavvik"))+
  scale_fill_manual(
    values = c('darkblue', 'lightblue'),
    labels = c('Effektivitet', 'Trivsel')
  )+
  geom_errorbar(
    aes(
      # x=nts,
      ymin = lower,
      ymax = upper),
      color = "black",
    position=position_dodge(.9),
    width=.4
  ) +
  # facet_grid(cols = vars(Type))+
  labs(y="P\u00E5virkning", x="St\u00F8ytype") + 
  scale_x_discrete(limit = nts)+
  theme_minimal() + 
  theme(legend.title = element_blank())
```

```{r saveNTconfInt, eval=F, include=F}
pdf(file = './figures/noiseTypeBarConfInt.pdf', width = 6,height = 4)
ggplot(ntDfCiLong, aes(x=nts, y = mean ,fill= Type)) + 
  geom_col(position="dodge") + 
  # scale_y_continuous(sec.axis = sec_axis(~ . /ntDf_maxMean*ntDf_maxSD , name = "Standardavvik"))+
  scale_fill_manual(
    values = c('darkblue', 'lightblue'),
    labels = c('Effektivitet', 'Trivsel')
  )+
  geom_errorbar(
    aes(
      # x=nts,
      ymin = lower,
      ymax = upper),
      color = "black",
    position=position_dodge(.9),
    width=.4
  ) +
  # facet_grid(cols = vars(Type))+
  labs(y="P\u00E5virkning", x="St\u00F8ytype") + 
  scale_x_discrete(limit = nts)+
  theme_minimal() + 
  theme(legend.title = element_blank())
dev.off()
```

```{r verificationNTconfInt, include=F,eval=F}
hist(d$byggestoy1)
CI(d$byggestoy1)
sd(d$byggestoy1)
mean(d$byggestoy1)
mean(d$byggestoy1) + 1.96*sd(d$byggestoy1)/sqrt(143)
```

## NT Count (Discarded)
```{r ntCount, options}
valueCount = function(x){
  return(c(sum(x==1), sum(x==2), sum(x==3), sum(x==4), sum(x==5)))
}

ntCount = data.frame(
  cbind(
    score=1:length(nts),
    apply(d[,ntTriv], 2, FUN=valueCount)
    )
)

ntCountLong <- ntCount|>
  pivot_longer(cols=-score, names_to='Type')

ggplot(ntCountLong, aes(x=score, y = value ,fill= Type)) + 
  geom_col(position="dodge") + 
  # scale_y_continuous(sec.axis = sec_axis(~ . /ntDf_maxMean*ntDf_maxSD , name = "Standardavvik"))+
  # scale_fill_manual(
  #   values = c('darkblue', 'lightblue'),
  #   labels = c('Effektivitet', 'Trivsel')
  # )+
  # facet_grid(cols = vars(Type))+
  labs(y="Antall", x="Svar") + 
  # scale_x_discrete(limit = nts)+
  theme_minimal() + 
  theme(legend.title = element_blank())

nTemp = length(d[,'personstoy1'])
CI(d[,'personstoy1'])
mTemp = mean(d[,'personstoy1'])
sTemp = sd(d[,'personstoy1'])
mTemp + qt(0.975,nTemp)*(sTemp/sqrt(nTemp))
```


# Kjel and Varme
Combine `kjel` and `varme` because of struggle with polygons in "building colored" map.
```{r kjelVarme, options}
nKjel = length(d[d$spmHvor=='kjel',1])
nVarme = length(d[d$spmHvor=='varme',1])
nKV = nKjel + nVarme
vkTriv = c(
  d[d$spmHvor=='kjel','spmTriv'],
  d[d$spmHvor=='varme','spmTriv']
)
vkEff = c(
  d[d$spmHvor=='kjel','spmEff'],
  d[d$spmHvor=='varme','spmEff']
)
mTriv = mean(vkTriv)
mEff = mean(vkEff)
sdTriv = sd(vkTriv)
sdEff = sd(vkEff)
kjVarme = cbind(mTriv,mEff,sdTriv,sdEff)
colnames(kjVarme) = c('meanTriv', 'meanEff', 'sdTriv', 'sdEff')
round(kjVarme,2)
```


# ANC and noise type correlation
We will look at noise types and how much they disturb when the individual uses ANC compared to when the individual is not.
```{r ANCNoisetype, options}
# ntTriv = colnames(d)[7:11]
# ntEff = colnames(d)[13:17]
ntColnames = c(
  'Konstruksjon', 'Person', 'Trafikk', 'Vifte', 'Annet')
ntRownames = c(
  'Trivsel m/ ANC', 'Trivsel u/ ANC', 'Effektivitet m/ ANC',
  'Effektivitet u/ ANC')
ntAnc = data.frame(
  matrix(
    ncol = length(ntTriv), nrow = 4,
    dimnames=list(ntRownames, ntColnames)
  )
)
ntAncSD = data.frame(
  matrix(
    ncol = length(ntTriv), nrow = 4,
    dimnames=list(ntRownames, ntColnames)
  )
)

# Find mean noise type disturbances
ancBool = d[,"spmTiltak_1"] == "mNc"
ntAnc[ntColnames] = rbind(
  apply(d[ancBool, ntTriv], FUN = mean, MARGIN = 2),
  apply(d[!ancBool, ntTriv], FUN = mean, MARGIN = 2),
  apply(d[ancBool,ntEff], FUN = mean, MARGIN = 2),
  apply(d[!ancBool,ntEff], FUN = mean, MARGIN = 2)
)
ntAncSD[ntColnames] = rbind(
  apply(d[ancBool, ntTriv], FUN = sd, MARGIN = 2),
  apply(d[!ancBool, ntTriv], FUN = sd, MARGIN = 2),
  apply(d[ancBool,ntEff], FUN = sd, MARGIN = 2),
  apply(d[!ancBool,ntEff], FUN = sd, MARGIN = 2)
)

# Barplot
reOrder = c(1,3,2,4)
cols = c("#b1fbf5", '#d5a7fb', '#59f7ea', '#be77f9')

# Mean
barplot(
  ((as.matrix(ntAnc[reOrder,]))), 
  col = cols[reOrder],
  border = "white",
  # main="Trivsel",
  ylab="P\u00E5virkning",
  beside = T,
  # las=2,
  ylim = c(0,5)
  # space=0.1
)
legend(
  "top",
  legend = ntRownames[reOrder],
  fill = cols[reOrder], bty = 'n')

# SD
ntAncSDrange = max(ntAncSD)
barplot(
  ((as.matrix(ntAncSD[reOrder,]))), 
  col = cols[reOrder],
  border = "white",
  # main="Trivsel",
  ylab="Standardavvik",
  beside = T,
  # las=2,
  ylim = c(0,ntAncSDrange*(1+0.3))
  # space=0.1
)
legend(
  "top",
  legend = ntRownames[reOrder],
  fill = cols[reOrder], bty = 'n')
```

```{r saveANCnoiseType, eval=F, include=F}
par(mar=defaultMar-c(0,0,1,0))

# Mean
pdf(file = './figures/histANCnoiseType.pdf', width = 6,height = 6)
barplot(
  ((as.matrix(ntAnc[reOrder,]))), 
  col = cols[reOrder],
  border = "white",
  # main="Trivsel",
  ylab="P\u00E5virkning",
  beside = T,
  # las=2,
  ylim = c(0,5)
  # space=0.1
)
legend(
  "top",
  legend = ntRownames[reOrder],
  fill = cols[reOrder], bty = 'n')
dev.off()

# SD
pdf(file = './figures/histANCnoiseTypeSD.pdf', width = 6,height = 6)
barplot(
  ((as.matrix(ntAncSD[reOrder,]))), 
  col = cols[reOrder],
  border = "white",
  # main="Trivsel",
  ylab="Standardavvik",
  beside = T,
  # las=2,
  ylim = c(0,ntAncSDrange*(1+0.3))
  # space=0.1
)
legend(
  "top",
  legend = ntRownames[reOrder],
  fill = cols[reOrder], bty = 'n')
dev.off()

par(mar=defaultMar)
```

# LME models
Here we look at the nested linear mixed models explained in the theory of the report.
```{r LMEsetup, options}
# Setup
trivNames = names(d)[6:11]
effNames = names(d)[12:17]
tiltakNames = names(d)[18:24]
dt = d[,c(trivNames, tiltakNames, 'spmHvor')]
de = d[,c(effNames, tiltakNames, 'spmHvor')]

tiltakBool = function(x){
  # Function for random effects
  return(x!="")
}
```

## Model selection
First we fit the models.
```{r lmeFit, options}
# Trivsel ###################################
# Complex model
totMod_Triv = lmer(
  spmTriv ~ -1 + byggestoy1 + personstoy1 + trafikk1 + vifte1 + annet1 + as.factor(spmHvor)
   + (1|spmTiltak_1) + (1|spmTiltak_2) + (1|spmTiltak_3) + (1|spmTiltak_4)
  + (1|spmTiltak_5) + (1|spmTiltak_6) + (1|spmTiltak_7)
  ,data = dt)

# without tiltak
ntWhereMod_Triv = lm(
  spmTriv ~ -1 + byggestoy1 + personstoy1 + trafikk1 + vifte1 + annet1 + as.factor(spmHvor)
  ,data = dt)


# Only noise types
ntMod_Triv = lm(
  spmTriv ~ -1 + byggestoy1 + personstoy1 + trafikk1 + vifte1 + annet1
  ,data = dt)

# Additional model with only location as explanatory variable
whereMod_Triv = lm(spmTriv ~ as.factor(spmHvor), data=dt)


# Effektivitet ############################
# Complex model
totMod_Eff = lmer(
  spmEff ~ -1 + byggestoy2 + personstoy2 + trafikk2+ vifte2 + annet2 + as.factor(spmHvor)
   + (1|spmTiltak_1) + (1|spmTiltak_2) + (1|spmTiltak_3) + (1|spmTiltak_4)
  + (1|spmTiltak_5) + (1|spmTiltak_6) + (1|spmTiltak_7)
  ,data = de)

# Without tiltak
ntWhereMod_Eff = lm(
  spmEff ~ -1 + byggestoy2 + personstoy2 + trafikk2 + vifte2 + annet2 + as.factor(spmHvor)
  ,data = de)

# Only noise type
ntMod_Eff = lm(
  spmEff ~ -1 + byggestoy2 + personstoy2 + trafikk2 + vifte2 + annet2
  ,data = de)

# Additional model with only location as explanatory variable
whereMod_Eff = lm(spmEff ~ as.factor(spmHvor), data=de)
```

Then we make model selection using `anova`. 
```{r modSel, options}
# Trivsel
anova(totMod_Triv, ntWhereMod_Triv, ntMod_Triv, test='LRT') # In report
# Effektivitet
anova(totMod_Eff, ntWhereMod_Eff, ntMod_Eff, test='LRT')
```

```{r removed1, include=F, eval=F}
# anova(whereMod_Triv, ntWhereMod_Triv, test="Chisq")
# anova(ntWhereMod_Triv, ntMod_Triv, test='LRT')
# anova(ntWhereMod_Triv, ntMod_Triv, test='Chisq')

# anova(whereMod_Eff, ntWhereMod_Eff, test="Chisq")

# summary(totMod_Triv)
# summary(whereMod_Triv)
# summary(whereMod_Eff)
```

## Model summaries trivsel
We skip showing the summaries of nested models without significant $\chi^2$ on level $\alpha=0.1$. That is, we skip the total model for `Triv`.  
```{r lmeSummaries, options}
# Trivsel
summary(ntMod_Triv)
summary(ntWhereMod_Triv)

# Effektivitet
summary(ntMod_Eff)
summary(ntWhereMod_Eff)
```

# Misc requests
These are some statistics requested by team members.

## Anc data (old)
```{r anc, options}
N = length(d[,1])
ancPercentage = length(d$spmTiltak_1[d$spmTiltak_1=='mNc']) / N
ancPercentage
```

## Noise Type in Elektro (Sindre)
Closely consider the distribution of noise type answers on the scale 1-5 and each noise type mean and CI for location `elektro`.
```{r elektroNTdistr, options}
elektro = c("elD","elB","elE","elA")
elTriv = d[d$spmHvor %in% elektro, c('spmHvor', ntTriv)]
elEff = d[d$spmHvor %in% elektro, c('spmHvor', ntEff)]

elLong <- elTriv |> pivot_longer(cols = -spmHvor, names_to = "nt", values_to = 'triv')
elEffLong = elEff |> pivot_longer(cols = -spmHvor, names_to = "nt", values_to = 'eff')
elLong$eff = elEffLong$eff

# Plot
plotColors = colorRampPalette(c("lightblue", 'purple4'))(5)
ggElektroTriv = ggplot(data=elLong, aes(x=triv, fill=nt)) +
  geom_bar(position = 'dodge')+
  scale_fill_manual(
    values = plotColors,
    labels = c('Annet', 'Konstruksjon', 'personst\U00F8y', 'trafikk', 'vifte')
  ) +
  labs(x="P\u00E5virkning", y="Antall svar")+
  scale_y_continuous(breaks = seq(0,20,10)) +
  theme_minimal() +
  # remove the vertical grid lines
  theme(panel.grid.major.x = element_blank(), legend.title = element_blank())
ggElektroTriv

ggElektroEff = ggplot(data=elLong, aes(x=eff, fill=nt)) +
  geom_bar(position = 'dodge')+
  scale_fill_manual(
    values = plotColors,
    labels = c('Annet', 'Konstruksjon', 'personst\U00F8y', 'trafikk', 'vifte')
  ) +
  labs(x="P\u00E5virkning", y="Antall svar")+
  theme_minimal() +
  # remove the vertical grid lines
  theme(panel.grid.major.x = element_blank(), legend.title = element_blank())
ggElektroEff

# ggplot(data=elLong, aes(x=nt))+
#   geom_bar(position="dodge", aes(y = triv), stat='summary', fun='mean') +
#   geom_bar(aes(y = eff), stat='summary', fun='mean') 
```

```{r saveElektroNTDistr, include=FALSE, eval=FALSE}
pdf(file = './figures/noiseTypeDistrElektroTriv.pdf', width = 5,height = 3)
ggElektroTriv
dev.off()

pdf(file = './figures/noiseTypeDistrElektroEff.pdf', width = 5,height = 3)
ggElektroEff
dev.off()

# ggElektroTrivNoLegend = ggElektroTriv
# ggElektroTrivNoLegend + guides(fill = FALSE)
pdf(file = './figures/noiseTypeDistrElektroTrivNoLegend.pdf', width = 4,height = 3)
ggElektroTriv + guides(fill = FALSE)
dev.off()
```



```{r ntElektroMean, options}
elData = data.frame(
  cbind(
    nts=1:length(nts),
    t(apply(d[d$spmHvor %in% elektro, ntEff], FUN = CI, MARGIN=2)),
    t(apply(d[d$spmHvor %in% elektro, ntTriv], FUN = CI, MARGIN=2))
    )
)

meansTemp = c('nts','mean','mean.1')
upperTemp = c('nts','upper','upper.1')
lowerTemp = c('nts','lower','lower.1')
dfMTemp = elData[,meansTemp]
dfUTemp = elData[,upperTemp]
dfLTemp = elData[,lowerTemp]

ntDfCiLong <- dfMTemp|>
  pivot_longer(cols=-nts, names_to='Type')
ntDfULong = dfUTemp |>
  pivot_longer(cols=-nts, names_to='Type')
ntDfLLong = dfLTemp |>
  pivot_longer(cols=-nts, names_to='Type')

ntDfCiLong$upper=ntDfULong$value
ntDfCiLong$lower=ntDfLLong$value

ntDfCiLong$nts=nts[ntDfCiLong$nts]
ntDfCiLong$Type=rep(c('eff','triv'), as.integer(length(nts)))
colnames(ntDfCiLong) = c('nts', 'Type', 'mean', 'upper', 'lower')

ggElektro = ggplot(ntDfCiLong, aes(x=nts, y = mean ,fill= Type)) + 
  geom_col(position="dodge") + 
  scale_fill_manual(
    values = c('darkblue', 'lightblue'),
    labels = c('Effektivitet', 'Trivsel')
  )+
  geom_errorbar(
    aes(
      # x=nts,
      ymin = lower,
      ymax = upper),
      color = "black",
    position=position_dodge(.9),
    width=.4
  ) +
  labs(y="P\u00E5virkning", x="St\u00F8ytype") + 
  scale_x_discrete(limit = nts)+
  theme_minimal() + 
  theme(legend.title = element_blank())
ggElektro
```

```{r saveNTelektro, options}
pdf(file = './figures/noiseTypeBarConfIntElektro.pdf', width = 6,height = 4)
ggElektro
dev.off()
```



